---
title: "Schelling's Segregation Model"
author: "David Garcia, 2024"
output: html_document
editor_options: 
  chunk_output_type: console
---


# Tasks:

1. Code the first simulation with reusable functions

2. Code Moran's I and encapsulate simulation in a loop

3. Analyze the role of tolerance


# 1. Code the first simulation with reusable functions

Initialize model settings (grid size, number of agents per color) and the tolerance parameter.  Also load (and install if necessary) a package for plotting matrices with colors, such as plot.matrix.
```{r initialize}
# install.packages("plot.matrix")
library(plot.matrix)
M = 20 #grid size
N = 196 #number of agents of each color
Tolerance = 0.5 #Tolerance threshold
```

Code a function to initialize the random locations of agents.
```{r initializeLocations}
initializeLocations <- function(M=20, N=196)
{
  #Your code here
}
```


Code a function to find the neighboring cells of a cell with coordinates x and y. Consider coordinates in a torus using the modulo operator (%%).

```{r neighbors}
neighbors <- function(x,y)
{
  #Your code here
}
```

Code a function to calculate the fraction of neighbors of a fixed color for all cells in the grid. If a cell has no neighbors of any color, initialize this fraction as NA.

```{r neighborFractions}
neighborFractions <- function(locs)
{
  #Your code here
}
```

Using the functions you coded above, execute a run of Schelling's model. If you use a matrix plot function, add a small delay (e.g. 0.1 seconds) to the simulation so you can see the visualization. Run the simulation until no more changes take place when checking the happiness of all agents.

```{r oneSim, eval=F}
#Your code here
```


# 2. Code Moran's I and encapsulate simulation in a loop

Code a function to calculate the spatial autocorrelation of a simulation with Moran's I.

```{r MoransI}
MoransI <- function(locations)
{
  #Your code here
}
```

Encapsulate the code to make a simulation of Schelling's model into one function. Do not visualize anything, take only the threshold as a parameter and return the resulting Moran's I after no more changes take place.

```{r Schelling}
Schelling <- function(Tolerance)
{
 #Your code here
}
```

# 3. Analyze the role of tolerance

Run a loop iterating over thresholds between 0.05 and 0.55 with increments of 0.1. Run 10 simulations for each threshold and record the resulting Moran's I. Plot Moran's I as a function of tolerance threshold. How high is the threshold when segregation starts to increase drastically?

```{r Analysis}
 #Your code here
```

